\documentclass[12pt]{scrartcl} % Font size

\input{structure.tex} 

\title{	
	\normalfont\normalsize
	\textsc{Analysis with an Introduction to Proof}\\ % Your university, school and/or department name(s)
	\vspace{25pt} % Whitespace
	\rule{\linewidth}{0.5pt}\\ % Thin top horizontal rule
	\vspace{20pt} % Whitespace
	{\huge Sets and Functions, Chapter Exercises}\\ % The assignment title
	\vspace{12pt} % Whitespace
	\rule{\linewidth}{2pt}\\ % Thick bottom horizontal rule
	\vspace{12pt} % Whitespace
}

\author{\LARGE L. J.} % Your name

\date{\normalsize\today} % Today's date (\today) or a custom date

\newtheorem*{theorem}{Theorem}

\begin{document}

\maketitle % Print the title

\section{Section 1}

\nsubsection{2}{Mark each statement True or False. Justify each answer.}
\begin{enumerate}[(a)]
	\item \emph{False}. Let $A = \{x \in \mathbb{R}: x \text{ is prime}\}$ and let $B = \{x \in \mathbb{R}: x \text{ is divisible by 6} \}$. Then $A \cap B = \emptyset$, but $A \neq \emptyset$ and $B \neq \emptyset$.
	\item \emph{True}. Let $A = \{1, 3, 5\}$ and let $B = \{2, 4, 6\}$. Then $A \cup B = \{1, 2, 3, 4, 5, 6\}$. If $x \in A \cup B$, then $x$ is equal to one of those values. If $x = 4$, then it is also true that $x \in B$. By the same logic more generally, if $x \in A \cup B$, then $x \in A$ or $x \in B$.
	\item \emph{False}. If $x \in A\setminus B$, then $x \in A$ \emph{and} $x \notin B$. Let $C = \{x: (x \in A) \land (x \in B)\}$. If $x \in C$, then if $x \in A\setminus B \stackrel{?}{\implies} (x \in A) \lor (x \notin B)$, it is also true $x \in A\setminus B$. But this is at odds with the definition of the relative complement of $B$ with respect to $A$.
	\item \emph{False}. It's fine to begin this way; the only nontrivial proof is one in which $S$ is nonempty.
\end{enumerate}

\nsubsection{4}{Let $A = \{2,4,6,8\}$, $B =\{6,8,10\}$, and $C = \{5,6,7,8\}$. Find the following sets.}
\begin{enumerate}[(a)]
	\item $\{6,8\}$
	\item $\{2,4,6,8,10\}$
	\item $\{2,4\}$
	\item $\{6,8\}$
	\item $\{10\}$
	\item $\{5,7,10\}$
	\item $\emptyset$
	\item $\{5,7\}$
\end{enumerate}

\nsubsection{6}{Let A and B be subsets of a universal set U. Simplify each of the following expressions.}
\begin{enumerate}[(a)]
	\item $U$
	\item $\emptyset$
	\item $A \cap B$
	\item $A \cup B$
	\item $A$
	\item $A$
\end{enumerate}

\nsubsection{8}{Let $S = \{\emptyset, \{\emptyset\}\}$. Determine whether each of the following is True or False. Explain your answers.}
\begin{enumerate}[(a)]
	\item \emph{True}. $\forall x \in \emptyset$, $x \in S$. Therefore, $\emptyset \subseteq S$.
	\item \emph{True}. $S$ is strictly enumerated, and $\emptyset$ is one of its elements.
	\item \emph{True}. $\{\emptyset\} \subseteq S$ because for all $x \in \{\emptyset\}$, $x \in S$. Unlike in (a), though, the proof is not vacuous; $\emptyset$ is an element of both sets.
	\item \emph{True}. $\{\emptyset\}$ is an element in $S$, as defined by enumeration. (To contrast (c) and (d), $\{\{\emptyset\}\}$ would also be a valid subset of $S$.)
\end{enumerate}
\nsubsection{10}{Fill in the blanks in the proof of the following theorem.}
\begin{theorem}
$A \subseteq B$ iff $A \cup B = B$.
\end{theorem}
\begin{proof}
Suppose that $A \subseteq B = B$. If $x \in A \cup B$, then $x \in A$ or $x \in B$.
Since $A \subseteq B$, in either case, we have $x \in B$. Thus $A \subseteq B$.
On the other hand, if $x \in B$, then $x \in A \cup B$, so $A \subseteq B$. Hence $A \cup B \ B$. \\
Conversely, suppose that $A \cup B = B$. If $x \in A$, then $x \in A \cup B$. But $A \cup B = B$, so $x \in B$. Thus $A \subseteq B$.
\end{proof}
\nsubsection{14}{Which statements below would enable one to conclude that $x \in A \cup B$?}
\begin{enumerate}[(a)]
	\item If $x \in A$ and $x \in B$, then $x \in A \cup B$.
	\item If $x \in A$ or $x \in B$, then $x \in A \cup B$.
	\item This statement tells us $A \subseteq B$, but does not let us conclude anything about $x$.
	\item If $x \notin A$, then $x \in B$ implies that if $x \notin B$, then $x \in A$. Since $x$ must always be either in $A$ or $B$, we can conclude that $x \in A \cup B$.
\end{enumerate}

\nsubsection{16}{Which statements below would enable one to conclude that $x \in A\setminus B$?}
\begin{enumerate}[(a)]
	\item If $x \in A$ and $x \notin B\setminus A$, then $x$ could be in $A \cap B$, so we cannot conclude that $x \in A \setminus B$.
	\item If $x \in A \cup B$ and $x \notin B$, then $(A \cup B) \cap U\setminus B$. Equivalently, $A \cap U\setminus B \cup B \cap U\setminus B$, and $A\setminus B \cup \emptyset$, or $A\setminus B$. Therefore we can conclude that $x \in A\setminus B$.
	\item If $x \in A \cup B$ and $x \notin A \cap B$, then $x \in A$ or $x \in B$. If $x \in A$, then $x \in A\setminus B$. But if $x \in B$, then $x \notin A\setminus B$. So we cannot conclude that $x \in A\setminus B$.
	\item If $x \in A$ and $x \notin A \cap B$, then $x \notin B$. Thus $x \in A\setminus B$
\end{enumerate}

\nsubsection{18}{Prove that the empty set is unique. That is, suppose that $A$ and $B$ are empty and prove that $A = B$.}
\begin{proof}
Suppose that $A$ and $B$ are empty sets. If $A = B$, then $A \subseteq B$ and $B \subseteq A$. \\
If $A \subseteq B$, then if $x \in A$, $x \in B$. By the definition of an empty set, for all $x$, $x \notin \emptyset$. Therefore, for all $x$, $x \notin A$ and $x \notin B$. By logical implication, then, if $x \in A$, $x \in B$. This implies that $A \subseteq B$. \\
On the other hand. if $x \in B$, then $x \in A$. By the same argument, we conclude $B \subseteq A$. So, we have $A \subseteq B$ and $B \subseteq A$. Thus $A = B$.
\end{proof}

\nsubsection{20}{Prove: $A \cap B$ and $A\setminus B$ are disjoint and $A = (A \cap B) \cup (A \setminus B)$.}
\begin{proof}
Assume the conclusion that $(A \cap B) \cap (A\setminus B) = \emptyset$. \\
If $x \in (A \cap B)$, then $x \in A$ and $x \in B$. But then $x \notin (A \setminus B)$, because if $x \in A \setminus B$, then $x \in A$ and $x \notin B$. In other words, if $x \in (A \cap B)$, $x \in (A \setminus (A \setminus B))$. So $(A \cap B) \nsubseteq (A \setminus B)$. \\
Conversely, if $x \in (A\setminus B)$, then $x \in A$ and $x \in B$. But then $x \notin (A \cap B)$. So $(A\setminus B) \nsubseteq (A \cap B)$. Thus, $(A \cap B) \cap (A\setminus B) = \emptyset$. Therefore $(A \cap B)$ and  $(A\setminus B)$ are disjoint. \\
In other words:
\begin{align}
(A \cap B) \cap (A\setminus B) \\
= (A \cap B) \cap (A \cap U\setminus B) \\
= (U\setminus A \cup U\setminus B) \cup (U\setminus A \cup B) \\
= (U\setminus A) \cup (U\setminus B \cup B) \\
= A \cap (B \cap U\setminus B) \\
= A \cap \emptyset \\
= \emptyset
\end{align}
\end{proof}
\begin{proof}
To prove  $A = (A \cap B) \cup (A \setminus B)$, first show that if $x \in (A \cap B) \cup (A\setminus B)$, then $x \in A$. \\
If $x \in (A \cap B) \cup (A\setminus B)$, then $x \in (A \cup A\setminus B)$ and $x \in (B \cup A\setminus B)$. Suppose $x \in (A \cap B) \cup (A\setminus B)$.
If $x \in A$, then $x \in (A \cup A\setminus B)$ and $x \in (B \cup A\setminus B)$, so $x \in (A \cap B) \cup (A\setminus B)$. \\If $x \in B$, then $x \in (B \cup A\setminus B)$ (equivalently, $(A \cup B)$), but $x \notin (A \cup A\setminus B)$ (equivalently, $A$). So by necessity, we have $x \in A$, and $(A \cap B) \cup (A\setminus B) = A$. \\

On the other hand, suppose $x \in A$. Then $x \in (A \cap B)$ or $x \in (A\setminus B)$:
\begin{align}
A\\
= A \cap U \\
= A \cap (B \cup U\setminus B) \\
= (A \cap B) \cup (A \cap U\setminus B) \\
= (A \cap B) \cup (A\setminus B)
\end{align}
\\Either way, $x \in (A \cap B) \cup (A \setminus B)$. So we have $A = (A \cap B) \cup (A \setminus B)$.

\end{proof}

\section{Section 2}

\nsubsection{2}{Mark each statement True or False. Justify each answer.}
\begin{enumerate}[(a)]
	\item \emph{True}. If $\mathscr{P}$ is a partition of $S$, then if $x$ and $y$ are in the same part of the partition, we can define $x R y$ as an equivalence relation such that the equivalence classes are the same as $\mathscr{P}$.
	\item \emph{False}. If $xRx$ for all $x \in S$, then $R$ is reflexive. But not all relations are reflexive.
	\item \emph{False}. A partition of $S$ can be defined by $\{E_x: x \in S \}$, where $E_x$ represents an equivalence class. An equivalence class is defined by $E_x = \{y \in S: yRx\}$.
	\item \emph{True}. By definition, a partition $\mathscr{P}$ consists of nonempty, disjoint sets such that for all $x \in S$, $x$ belongs to some subset of $\mathscr{P}$.
\end{enumerate}

\nsubsection{4}{Show that $\{a\} \times \{a\} = \{\{\{a\}\}\}$.}
$(a,b)$  is the set whose members are $\{a\}$ and $\{a,b\}$.
	\begin{align*}
	(a,b) &= \{\{a\}, \{a,b\}\}
	\end{align*}
	And the product of two sets $A$ and $B$ is
	\begin{align*}
	A \times B \ \{(a,b) &: a \in A \text{and} b \in B \}
	\end{align*}		
	Then,
	\begin{align*}
	\{a\} \times \{a\} &= \{(a,a)\} \\
	&= \{\{ \{a\}, \{a, a\} \}\}.
	\end{align*}
	Since $a = a$, $\{a, b\} = \{a\}$, and the above simplifies to
	\begin{align*}
	&= \{\{\{a\} \{a\}\}\} \\
	&= \{\{\{a\}\}\}.
	\end{align*}

\nsubsection{6}{Let $A$ be be any set and let $B=\emptyset$. What can you conclude about $A \times B$?}
	Let $A$ be any set, and let $B$ be $\emptyset$. 
	By definition, 
	\begin{align*}
	A\times B = \{(a,b): a \in A \text{ and } b \in B \}
	\end{align*}
	
and
	\begin{align*}
	\emptyset = \{x: &\forall x, x\notin \emptyset\}.
	\end{align*}
	
	
	Then if $B = \emptyset$, the product $A\times B$ requires that all the ordered pairs in the product have a first element in $A$ and a second element in $\emptyset$. But now, $\forall b, b \notin B$, so there are no ordered pairs that satisfy the definition of the product of $A \times B$.  \\
	Therefore, $A \times B= \emptyset$.
	
	
\nsubsection{8}{Let $A = \{1, 2\}$.}
\begin{enumerate}[(a)]
\item \textbf{How many elements are in the set $A \times A$?}
$\mathopen|A \times A \mathclose| = 4$.
\item \textbf{How many possible relations are there on the set $A$?}
A relation on $A$ is any subset $R$ of $A \times A$ ($R \subset A \times A$); any element can be in or out of the relation. So, what is the power set of $A$?
$$ 2^{\mathopen|A \times A \mathclose|} = 2^4 = 16 $$
\item \textbf{How many possible relations are there on the set $\{1, 2, 3\}$?}
For $A$, we had $\mathopen|A \mathclose| = 2$. Let $S$ be a generic set, $n = \mathopen| S \mathclose|$ be the number of elements in the set, and $\mathopen| R \mathclose|$ be the total number of possible relations on $S$. Then for $S_n$, we have $\mathopen| R \mathclose| = 2^{n^2}$. If $S = \{1, 2, 3\}$, then $n = 3$ and we have
$$ \mathopen| R \mathclose| = 2^{(3^2)} = 2^9 = 512 $$ possible relations.
\end{enumerate}

\nsubsection{10}{Prove or give a counterexample.}
\begin{enumerate}[(a)]
\item Let $A = \{2, 3\}$ and $B = \{0, -5\}$.
Then $$ A \times B = \{(2,0), (2,-5), (3,0), (3,5)\} $$ and 
$$ B \times A = \{(0,2), (0,3), (-5,2), (-5,3)\}. $$ Therefore $$ A \times B \neq B \times A. $$
\item $(A \cup B) \times C = (A \times C) \cup (B \times C)$. \\
Let $$ A = \{x: x \in [0, 3)\} $$
$$ B = \{x: x \in (2, 4]\} $$
$$ C = \{x: x \in [-2, 2]\}. $$

Then $$ (A \cup B) \times C = \{(x,y): x \in [0, 4] \text{ and } y \in [-2, 2] $$

We have:
$$ A \times C = \{(x, y): x \in [0,3) \text{ and } y \in [-2,2]\} $$
$$ B \times C \ \{(x, y): x \in (2, 4] \text{ and } y \in [-2, 2]\} $$

and 

$$ (A \times C) \cup (B \times C) = \{(x, y): x \in [0, 4] \text{ and } y \in [-2, 2] \}.$$
\item TK
\item TK
\end{enumerate}
	
\nsubsection{12}{Find examples of relations with the following properties.}
Let $S = \{p, q, r\}$. Then define the following relations $R$ on $S$:
\begin{enumerate}[(a)]
\item \textbf{$R$ is reflexive, but not symmetric or transitive.} \\ $R = \{(p,p),(q,q),(r,r,)\}$
\item \textbf{$R$ is symmetric, but not reflexive or transitive.} \\ $R = \{(p,q), (q,p), (r,q), (q,r)\}$
\item \textbf{$R$ is transitive, but not symmetric or reflexive.} \\ $R = \{(p,q), (q,r), (p,r)\}$
\item \textbf{$R$ is reflexive and symmetric, but not transitive.} \\ $R = \{(p,p), (q,q), (r,r), (p,q), (q,p), (q,r), (r,q)\}$
\item \textbf{$R$ is reflexive and transitive, but not symmetric.} \\ $R = \{(p,p), (q,q), (r,r), (p,q), (q,r), (p,r)\}$
\item \textbf{$R$ is symmetric and transitive, but not reflexive.} \\ $R = \{(p, q), (q, p), (q, r), (r, q), (p, r), (r, p)\}$
\end{enumerate}

\nsubsection{14}{Let $S = \mathbb{R} \times \mathbb{R}$. Verify that the relation $(a,b)R(c,d)$ iff $a + d = b + c$ is an equivalence relation. Describe the equivalence class $E_{(7,3)}$.}

Rearrange the relation as:
\begin{align*}
a+d &= b+c \\
a+d-c &= b \\
a-c &= b-d
\end{align*}

The relation is reflexive: $(a,b)R(a,b)$ gives
$$ a-a = b-b $$
$$ 0 = 0 $$

The relation is symmetric:
\begin{align*}
(c,d) &R (a,b) \\
c-a &= d-b \\
-c+a &= -d+b  \\
a-c &= b-d
\end{align*}

The relation is transitive:
$$(a,b)R(c,d) \land (c,d)R(m,n) \implies (a,b)R(m,n)$$

\begin{align*}
TK TK TK TK
\end{align*}

The equivalence class $E_{(7,3)} = \{d \in S: d = c-4\}$:
\begin{align*}
7 + d &= 3 + c \\
d &= c - 4
\end{align*}
It describes a linear relationship.

\nsubsection{16}{Let $S = \mathbb{R} \times \mathbb{R}$. Define the equivalence relation $R$ on $S$ by $(a,b)R(c,d)$ iff $b-3a = d - 3c$.}

\begin{enumerate}[(a)]

\item We can rearrange the relation as:

\begin{align*}
b-d &= 3a - 3c \\
b-d &= 3(a-c) \\
\end{align*}

So, $R$ partitions $S$ into parallel lines of the form $y = 3x$ $(+ k)$.

\item $E_{(2,5)}$ would be the part of the partition containing the point $(2,5)$, which we obtain by substituting it for one-half of the ordered pairs:

\begin{align*}
b-3a &= 5 - 3(2) \\
b-3a &= -1
x &= 3a -1
\end{align*}

So the equivalence class is described by the linear relationship $y = 3x -1$.
\end{enumerate}

\nsubsection{18}{Let $R$ be the relation $\{(1,1), (2,2), (3,1), (3,3)\}$ on the set $\{1,2,3\}$. If $R$ is an equivalence relation, list the pieces of the partition determined by $R$, and if not, state why.}

$R$ is not an equivalence relation because it is not symmetric. $(3,1) \in R$ but $(1,3) \notin R$.

\nsubsection{20}{Let $S = \{a, b, c, d\}$ and let $P = \{\{a\}, \{b, c, d\}\}$. Describe the equivalence relation $R$ on $S$ determined by $P$ by listing the ordered pairs in the relation.}

We see that $a$ is in a partition alone, but $b, c$ and $d$ are in a relation with each other, so the relation consists of the pairs

\begin{align*}
\{ &(a,a), \\
&(b,b), (c,c), (d,d), \\
&(b,c), (b,d), \\
&(c,b), (c,d), \\
&(d,c), (d,b) \}.
\end{align*}

\end{document}